Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-23 doc/asplos13/evaluation.tex
split UTS scaling and 1-16 sections. Trim the XMT retries discussion down

We ran UTS-mem on Grappa and the XMT with a binomial and two geometric
100M-vertex trees (T3L,T1L,T2L), a geometric 1.6B-vertex tree (T1XL),
and a geometric 4.2B-vertex tree (T1XXL). Both systems do relatively
poorly on the binomial tree because it is very deep and a traversal
produces too little parallelism. Figure~\ref{fig:grappa-xmt-uts} shows
the performance in terms of number of vertices visited per second versus
number of compute nodes. The Grappa results shown are for 5 cores per
node and the best parameter values from a search over runtime and
application parameters. Grappa with \checkme{16} machines is faster than
the entire XMT of 128 processors. Grappa achieves \checkme{188Mvert/s}
with \checkme{128 machines} and the XMT achieves only 50Mvert/s,
plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
scaling up through 128---the maximum we could use---although scaling
declines because of the unscalability of our aggregation mechanism as
number of network endpoints increases. Using 6 cores per node gave
Grappa up to a 20\% benefit at each data point, but Grappa could not run
well beyond 90 machines with this core configuration because of the
current network aggregator design.

@@

Despite our efforts to tune the UTS implementation specific to the 
XMT, performance does not scale well with increasing processor count,
flattening out around sixty processors.  When we increase the size of
the tree from 100M to 4.2B, we find that performance does not improve, suggesting that performance is not limited by task parallelism.   Cray's performance tools show that an increasing number of memory retry operations are generated by code within the runtime system.  Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying responses at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

@@

>>>>

%Using 6 cores per node gave
%Grappa up to a 20\% benefit at each data point, but Grappa could not run
%well beyond 90 machines with this core configuration because of the
%current network aggregator designa.

We ran UTS-mem on Grappa and the XMT with a binomial and two geometric
100M-vertex trees (T3L,T1L,T2L), a geometric 1.6B-vertex tree (T1XL),
and a geometric 4.2B-vertex tree (T1XXL). Both systems do relatively
poorly on the binomial tree because it is very deep and a traversal
produces too little parallelism. Figure~\ref{fig:grappa-xmt-uts} shows
the performance in terms of number of vertices visited per second versus
number of compute nodes. The Grappa results shown are for 6 cores per
node and the best parameter values from a search over runtime and
application parameters.

@@

@@

\subsection{Scaling}
\paragraph{UTS-Mem}
We ran UTS-Mem on Grappa and the XMT with a geometric 1.6B-vertex tree
(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
nodes---the maximum we had available for each. Grappa results are for 5 cores per node. Grappa with 20 machines is faster than the entire XMT of 128 processors.
Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
scaling up through 128 nodes, although scaling
declines because of the unscalability of our aggregation mechanism as
number of network endpoints increases.

Despite our efforts to tune the UTS implementation specific to the 
XMT, performance does not scale well with increasing processor count,
flattening out around 60 processors.  When we increase the size of
the tree from 100M to 4.2B, we find that performance does not improve,
suggesting that performance is not limited by task parallelism.
Cray's performance tools show an increasing number of memory
retry operations for failed synchronization operations generated by
the runtime, which create network contention.
%Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying responses at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.
